---
title: "KaggleGroup:House Price Prediction"
author: "Dheeraj Yata"
date: "2023-12-01"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
# This code chunk sets global options.

knitr::opts_chunk$set(echo = TRUE, message= F, warning = F)

# message= F and  warning = F suppress verbose information getting printed in 
# the compiled document.

# echo = T ensures that your code is visible.

# Load packages
library(tidyverse)
library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
```

## 1.Introduction & Project Goal with Aims:

to be changed

This is a project in IS-6489, where we embark on an exciting journey in the world of data science by participating in the Kaggle competition titled "House Prices: Advanced Regression Techniques." This competition, situated within Kaggle's playground, offers an invaluable opportunity to apply our data analytics skills to real-world data and challenges.

Our primary goal for this project is to develop a robust predictive model for housing prices in Ames, Iowa. The competition data, thoughtfully split into a training set and a test set, presents us with an intriguing challenge: to construct a model that accurately estimates house prices while maintaining simplicity. The minimum benchmark for our model's estimated out-of-sample R² is set at 0.75, reflecting our aim to achieve a balance between predictive performance and model parsimony.

## 2.Data import and inspection

```{r}
#importing the data
test <- read_csv("test.csv")
train <- read_csv("train.csv")
submit_example <- read_csv("sample_submission.csv")
# Train data inspection and displaying top data
head(train)
```

```{r}
#Getting rid of the IDs but keeping the test IDs in a vector. These are needed to compose the submission file
test_labels <- test$Id
test$Id <- NULL
train$Id <- NULL
head(train)
```

Without the Id’s, the dataframe consists of 79 predictors and our response variable SalePrice.

```{r}
#combining two data sets for better understanding
test$SalePrice <- NA
all <- rbind(train, test)
dim(all)
```


```{r}
#test data Inspection
head(test)
```
The test dataset also consist of character and integer variables. Most of the character variables are actually (ordinal) factors. In total, there are 80 columns/variables,the missing response variable is (SalePrice) it is to be predicted by the end of the modeling. Above is a glimpse of the variables

```{r}
# example of submitting predictions
head(submit_example)
```
It has two columns one with Id's of individual houses and other the predicted values of sales price.

## 3. Exploring  some of the most important variables.
### 3.1 The response variable : Sales price
```{r}

library(scales) 

#Histogram plot of SalePrice of train data set
ggplot(data=train[!is.na(train$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="blue", binwidth = 10000) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

we can see, the sale prices are right skewed. This was expected as few people can afford very expensive houses.

```{r}
# summary of saleprice in train data
summary(train$SalePrice)
```
### 3.2 The most important numeric predictors
```{r pressure, echo=FALSE}
numericVars <- which(sapply(train, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
cat('There are', length(numericVars), 'numeric variables')
```
### 3.2.1 Correlations with SalePrice

```{r}
library(corrplot)

train_numVar <- train[, numericVars]
cor_numVar <- cor(train_numVar, use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")
```
Lets explore more on the relation between SalePrice and the two predictors with the highest correlation with SalePrice; Overall Quality and the ‘Above Grade’ Living Area.

### 3.2.2 Overall Quality

`OverallQual` is a categorical feature (quality split into 10 categories), but is encoded as numeric.   If it is non-linearly related to the target then encoding it as a factor would work best.  If the relationship is linear then numeric is fine.

```{r}
#checking the relationship between OverallQual and SalePrice
ggplot(data=train[!is.na(train$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot(col='blue') + labs(x='Overall Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)

```

### 3.2.3 Above Ground Living Area (square feet)

The numeric variable with the second highest correlation with SalesPrice is the Above Grade Living Area. This make a lot of sense; big houses are generally more expensive.

```{r}
library(ggrepel)

ggplot(data=train[!is.na(train$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_text_repel(aes(label = ifelse(train$GrLivArea[!is.na(train$SalePrice)]>4500, rownames(train), '')))
```
## 4.Missing data, label encoding, and factorizing variables
### 4.1 Completeness of the data

First lets  see which variables contain missing values.

```{r}
#exploring all missing values
NAcol <- which(colSums(is.na(all)) > 0)
sort(colSums(sapply(all[NAcol], is.na)), decreasing = TRUE)
```
```{r}
cat('There are', length(NAcol), 'columns with missing values')
```
Of course, the 1459 NAs in SalePrice match the size of the test set perfectly. This means that we have to fix NAs in 34 predictor variables.

## 4.2 Imputing missing data

Here We are going to fix the 34 predictors that contains missing values. We will go through them working my way down from most NAs until I have fixed them all. If we stumble upon a variable that actually forms a group with other variables, we will also deal with them as a group. For instance, there are multiple variables that relate to Pool, Garage, and Basement.

Besides making sure that the NAs are taken care off, we have also converted character variables into ordinal integers if there is clear ordinality, or into factors if levels are categories without ordinality.

`Pool Quality and the PoolArea variable`

The PoolQC is the variable with most NAs. The description is as follows:

PoolQC: Pool quality

   Ex   Excellent
   Gd   Good
   TA   Average/Typical
   Fa   Fair
   NA   No Pool

So, it is obvious that we need to just assign ‘No Pool’ to the NAs. Also, the high number of NAs makes sense as normally only a small proportion of houses have a pool.

```{r}
all$PoolQC[is.na(all$PoolQC)] <- 'None'
```


```{r}
# Encode this variable as the values are ordinal
Qualities <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
```

```{r}
library(plyr)
all$PoolQC<-as.integer(revalue(all$PoolQC, Qualities))
table(all$PoolQC)
```
However, there is a second variable that relates to Pools. This is the PoolArea variable (in square feet). As you can see below, there are 3 houses without PoolQC. First, We checked if there was a clear relation between the PoolArea and the PoolQC. As We did not see a clear relation (bigger of smaller pools with better PoolQC), We are going to impute PoolQC values based on the Overall Quality of the houses (which is not very high for those 3 houses).

```{r}
all[all$PoolArea>0 & all$PoolQC==0, c('PoolArea', 'PoolQC', 'OverallQual')]
```

```{r}
all$PoolQC[2421] <- 2
all$PoolQC[2504] <- 3
all$PoolQC[2600] <- 2
```

`Alley` Within Alley, there are 2721 NAs. As the values are not ordinal, I will convert Alley into a factor. Values:
   Grvl Gravel
   Pave Paved
   NA   No alley access

```{r}
all$Alley[is.na(all$Alley)] <- 'None'
all$Alley <- as.factor(all$Alley)

ggplot(all[!is.na(all$SalePrice),], aes(x=Alley, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue')+
        scale_y_continuous(breaks= seq(0, 200000, by=50000), labels = comma)
```
`Miscellaneous` feature not covered in other categories

Within Miscellaneous Feature, there are 2814 NAs. As the values are not ordinal, I will convert MiscFeature into a factor. Values:

   Elev Elevator
   Gar2 2nd Garage (if not described in garage section)
   Othr Other
   Shed Shed (over 100 SF)
   TenC Tennis Court
   NA   None

```{r}
all$MiscFeature[is.na(all$MiscFeature)] <- 'None'
all$MiscFeature <- as.factor(all$MiscFeature)

ggplot(all[!is.na(all$SalePrice),], aes(x=MiscFeature, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..))
```
When looking at the frequencies, the variable seems irrelevant to me. Having a shed probably means ‘no Garage’, which would explain the lower sales price for Shed. Also, while it makes a lot of sense that a house with a Tennis court is expensive, there is only one house with a tennis court in the training set.

`Fence quality`Within Fence, there are 2348 NAs. The values seem to be ordinal. Values:

   GdPrv    Good Privacy
   MnPrv    Minimum Privacy
   GdWo Good Wood
   MnWw Minimum Wood/Wire
   NA   No Fence
   
```{r}
all$Fence[is.na(all$Fence)] <- 'None'
table(all$Fence)
```
Our conclusion is that the values do not seem ordinal (no fence is best). Therefore, we converted Fence into a factor.

```{r}
all$Fence <- as.factor(all$Fence)
```

`Fireplace quality`The number of NAs in FireplaceQu matches the number of houses with 0 fireplaces. This means that We can safely replace the NAs in FireplaceQu with ‘no fireplace’. The values are ordinal, and We can use the Qualities vector that We have already created for the Pool Quality. Values:

   Ex   Excellent - Exceptional Masonry Fireplace
   Gd   Good - Masonry Fireplace in main level
   TA   Average - Prefabricated Fireplace in main living area or Masonry Fireplace in basement
   Fa   Fair - Prefabricated Fireplace in basement
   Po   Poor - Ben Franklin Stove
   NA   No Fireplace

```{r}
all$FireplaceQu[is.na(all$FireplaceQu)] <- 'None'
all$FireplaceQu<-as.integer(revalue(all$FireplaceQu, Qualities))
table(all$FireplaceQu)
```
`Number of fireplaces`Fireplaces is an integer variable, and there are no missing values.

```{r}
table(all$Fireplaces)
```
`LotFrontage`: Linear feet of street connected to property

486 NAs. The most reasonable imputation seems to take the median per neigborhood

```{r}
ggplot(all[!is.na(all$LotFrontage),], aes(x=as.factor(Neighborhood), y=LotFrontage)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
for (i in 1:nrow(all)){
        if(is.na(all$LotFrontage[i])){
               all$LotFrontage[i] <- as.integer(median(all$LotFrontage[all$Neighborhood==all$Neighborhood[i]], na.rm=TRUE)) 
        }
}
```


`LotShape`: General shape of property.No NAs. Values seem ordinal (Regular=best)

   Reg  Regular 
   IR1  Slightly irregular
   IR2  Moderately Irregular
   IR3  Irregular

```{r}
all$LotShape<-as.integer(revalue(all$LotShape, c('IR3'=0, 'IR2'=1, 'IR1'=2, 'Reg'=3)))
table(all$LotShape)
```
`LotConfig`: Lot configuration

No NAs. The values seemed possibly ordinal to me, but the visualization does not show this. Therefore, We will convert the variable into a factor.

   Inside   Inside lot
   Corner   Corner lot
   CulDSac  Cul-de-sac
   FR2  Frontage on 2 sides of property
   FR3  Frontage on 3 sides of property
   
```{r}
ggplot(all[!is.na(all$SalePrice),], aes(x=as.factor(LotConfig), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue')+
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..))
```

```{r}
all$LotConfig <- as.factor(all$LotConfig)
table(all$LotConfig)
```
`Garage variables`
Altogether, there are 7 variables related to garages.Two of those have one NA (GarageCars and GarageArea), one has 157 NAs (GarageType), 4 variables have 159 NAs.

First of all, We replaced all 159 missing GarageYrBlt: Year garage was built values with the values in YearBuilt (this is similar to YearRemodAdd, which also defaults to YearBuilt if no remodeling or additions).

```{r}
all$GarageYrBlt[is.na(all$GarageYrBlt)] <- all$YearBuilt[is.na(all$GarageYrBlt)]
```

As NAs mean ‘No Garage’ for character variables, We want to find out where the differences between the 157 NA GarageType and the other 3 character variables with 159 NAs come from.

```{r}
#check if all 157 NAs are the same observations among the variables with 157/159 NAs
length(which(is.na(all$GarageType) & is.na(all$GarageFinish) & is.na(all$GarageCond) & is.na(all$GarageQual)))
```


```{r}
#Find the 2 additional NAs
kable(all[!is.na(all$GarageType) & is.na(all$GarageFinish), c('GarageCars', 'GarageArea', 'GarageType', 'GarageCond', 'GarageQual', 'GarageFinish')])
```

The 157 NAs within GarageType all turn out to be NA in GarageCondition, GarageQuality, and GarageFinish as well. The differences are found in houses 2127 and 2577. As you can see, house 2127 actually does seem to have a Garage and house 2577 does not. Therefore, there should be 158 houses without a Garage. To fix house 2127, We have imputated the most common values (modes) for GarageCond, GarageQual, and GarageFinish.

```{r}
#Imputing modes.
all$GarageCond[2127] <- names(sort(-table(all$GarageCond)))[1]
all$GarageQual[2127] <- names(sort(-table(all$GarageQual)))[1]
all$GarageFinish[2127] <- names(sort(-table(all$GarageFinish)))[1]
```

```{r}
#display "fixed" house
kable(all[2127, c('GarageYrBlt', 'GarageCars', 'GarageArea', 'GarageType', 'GarageCond', 'GarageQual', 'GarageFinish')])
```

`GarageCars and GarageArea`: Size of garage in car capacity and Size of garage in square
Both have 1 NA. As you can see above, it is house 2577 for both variables. The problem probably occured as the GarageType for this house is “detached”, while all other Garage-variables seem to indicate that this house has no Garage.

```{r}
#fixing 3 values for house 2577
all$GarageCars[2577] <- 0
all$GarageArea[2577] <- 0
all$GarageType[2577] <- NA
```

```{r}
#check if NAs of the character variables are now all 158
length(which(is.na(all$GarageType) & is.na(all$GarageFinish) & is.na(all$GarageCond) & is.na(all$GarageQual)))
```
Now, the 4 character variables related to garage all have the same set of 158 NAs, which correspond to ‘No Garage’. We have fixed all of them.

`GarageType: Garage location`The values do not seem ordinal, so I will convert into a factor.

   2Types   More than one type of garage
   Attchd   Attached to home
   Basment  Basement Garage
   BuiltIn  Built-In (Garage part of house - typically has room above garage)
   CarPort  Car Port
   Detchd   Detached from home
   NA   No Garage
   
```{r}
all$GarageType[is.na(all$GarageType)] <- 'No Garage'
all$GarageType <- as.factor(all$GarageType)
table(all$GarageType)
```
`GarageFinish`: Interior finish of the garage. The values are ordinal.

   Fin  Finished
   RFn  Rough Finished  
   Unf  Unfinished
   NA   No Garage  
   
```{r}
all$GarageFinish[is.na(all$GarageFinish)] <- 'None'
Finish <- c('None'=0, 'Unf'=1, 'RFn'=2, 'Fin'=3)
```

```{r}
all$GarageFinish<-as.integer(revalue(all$GarageFinish, Finish))
table(all$GarageFinish)
```

`GarageQual`: Garage quality

Another variable than can be made ordinal with the Qualities vector.

   Ex   Excellent
   Gd   Good
   TA   Typical/Average
   Fa   Fair
   Po   Poor
   NA   No Garage
```{r}
all$GarageQual[is.na(all$GarageQual)] <- 'None'
all$GarageQual<-as.integer(revalue(all$GarageQual, Qualities))
table(all$GarageQual)
```
`GarageCond`: Garage condition. Another variable than can be made ordinal with the Qualities vector.

   Ex   Excellent
   Gd   Good
   TA   Typical/Average
   Fa   Fair
   Po   Poor
   NA   No Garage
   
```{r}
all$GarageCond[is.na(all$GarageCond)] <- 'None'
all$GarageCond<-as.integer(revalue(all$GarageCond, Qualities))
table(all$GarageCond)
```
`basement variables`Altogether, there are 11 variables that relate to the Basement of a house. Five of those have 79-82 NAs, six have one or two NAs.

```{r}
#check if all 79 NAs are the same observations among the variables with 80+ NAs
length(which(is.na(all$BsmtQual) & is.na(all$BsmtCond) & is.na(all$BsmtExposure) & is.na(all$BsmtFinType1) & is.na(all$BsmtFinType2)))
```

```{r}
#Find the additional NAs; BsmtFinType1 is the one with 79 NAs
all[!is.na(all$BsmtFinType1) & (is.na(all$BsmtCond)|is.na(all$BsmtQual)|is.na(all$BsmtExposure)|is.na(all$BsmtFinType2)), c('BsmtQual', 'BsmtCond', 'BsmtExposure', 'BsmtFinType1', 'BsmtFinType2')]
```
So altogether, it seems as if there are 79 houses without a basement, because the basement variables of the other houses with missing values are all 80% complete (missing 1 out of 5 values). Hence We imputed the modes to fix those 9 houses.

```{r}
#Imputing modes.
all$BsmtFinType2[333] <- names(sort(-table(all$BsmtFinType2)))[1]
all$BsmtExposure[c(949, 1488, 2349)] <- names(sort(-table(all$BsmtExposure)))[1]
all$BsmtCond[c(2041, 2186, 2525)] <- names(sort(-table(all$BsmtCond)))[1]
all$BsmtQual[c(2218, 2219)] <- names(sort(-table(all$BsmtQual)))[1]
```

Now that the 5 variables considered agree upon 79 houses with ‘no basement’.
BsmtQual: Evaluates the height of the basement
A variable than can be made ordinal with the Qualities vector.

   Ex   Excellent (100+ inches) 
   Gd   Good (90-99 inches)
   TA   Typical (80-89 inches)
   Fa   Fair (70-79 inches)
   Po   Poor (<70 inches
   NA   No Basement
   
```{r}
all$BsmtQual[is.na(all$BsmtQual)] <- 'None'
all$BsmtQual<-as.integer(revalue(all$BsmtQual, Qualities))
table(all$BsmtQual)
```
`BsmtCond`: Evaluates the general condition of the basement. A variable than can be made ordinal with the Qualities vector.

   Ex   Excellent
   Gd   Good
   TA   Typical - slight dampness allowed
   Fa   Fair - dampness or some cracking or settling
   Po   Poor - Severe cracking, settling, or wetness
   NA   No Basement
   
```{r}
all$BsmtCond[is.na(all$BsmtCond)] <- 'None'
all$BsmtCond<-as.integer(revalue(all$BsmtCond, Qualities))
table(all$BsmtCond)
```

`BsmtExposure`: Refers to walkout or garden level walls. A variable than can be made ordinal.

   Gd   Good Exposure
   Av   Average Exposure (split levels or foyers typically score average or above)  
   Mn   Mimimum Exposure
   No   No Exposure
   NA   No Basement
   
```{r}
all$BsmtExposure[is.na(all$BsmtExposure)] <- 'None'
Exposure <- c('None'=0, 'No'=1, 'Mn'=2, 'Av'=3, 'Gd'=4)
```

```{r}
all$BsmtExposure<-as.integer(revalue(all$BsmtExposure, Exposure))
table(all$BsmtExposure)
```

`BsmtFinType1`: Rating of basement finished area

A variable than can be made ordinal.

   GLQ  Good Living Quarters
   ALQ  Average Living Quarters
   BLQ  Below Average Living Quarters   
   Rec  Average Rec Room
   LwQ  Low Quality
   Unf  Unfinshed
   NA   No Basement
 
```{r}
all$BsmtFinType1[is.na(all$BsmtFinType1)] <- 'None'
FinType <- c('None'=0, 'Unf'=1, 'LwQ'=2, 'Rec'=3, 'BLQ'=4, 'ALQ'=5, 'GLQ'=6)

```
   
```{r}
all$BsmtFinType1<-as.integer(revalue(all$BsmtFinType1, FinType))
table(all$BsmtFinType1)
```
`BsmtFinType2`: Rating of basement finished area (if multiple types)

A variable than can be made ordinal with the FinType vector.

   GLQ  Good Living Quarters
   ALQ  Average Living Quarters
   BLQ  Below Average Living Quarters   
   Rec  Average Rec Room
   LwQ  Low Quality
   Unf  Unfinshed
   NA   No Basement
   
```{r}
all$BsmtFinType2[is.na(all$BsmtFinType2)] <- 'None'
FinType <- c('None'=0, 'Unf'=1, 'LwQ'=2, 'Rec'=3, 'BLQ'=4, 'ALQ'=5, 'GLQ'=6)
```


```{r}
all$BsmtFinType2<-as.integer(revalue(all$BsmtFinType2, FinType))
table(all$BsmtFinType2)
```
Remaining Basement variabes with just a few NAs. We still have to deal with those 6 variables that have 1 or 2 NAs.

```{r}
#display remaining NAs. Using BsmtQual as a reference for the 79 houses without basement agreed upon earlier
all[(is.na(all$BsmtFullBath)|is.na(all$BsmtHalfBath)|is.na(all$BsmtFinSF1)|is.na(all$BsmtFinSF2)|is.na(all$BsmtUnfSF)|is.na(all$TotalBsmtSF)), c('BsmtQual', 'BsmtFullBath', 'BsmtHalfBath', 'BsmtFinSF1', 'BsmtFinSF2', 'BsmtUnfSF', 'TotalBsmtSF')]
```
`BsmtFullBath`: Basement full bathrooms. An integer variable.
```{r}
all$BsmtFullBath[is.na(all$BsmtFullBath)] <-0
table(all$BsmtFullBath)
```
`BsmtHalfBath`: Basement half bathrooms. An integer variable.
```{r}
all$BsmtHalfBath[is.na(all$BsmtHalfBath)] <-0
table(all$BsmtHalfBath)
```
`BsmtFinSF1`: Type 1 finished square feet. An integer variable.
```{r}
all$BsmtFinSF1[is.na(all$BsmtFinSF1)] <-0

```

`BsmtFinSF2`: Type 2 finished square feet.An integer variable.
```{r}
all$BsmtFinSF2[is.na(all$BsmtFinSF2)] <-0
```


`BsmtUnfSF`: Unfinished square feet of basement area.An integer variable.
```{r}
all$BsmtUnfSF[is.na(all$BsmtUnfSF)] <-0
```

`TotalBsmtSF`: Total square feet of basement area. An integer variable.
```{r}
all$TotalBsmtSF[is.na(all$TotalBsmtSF)] <-0
```

`Masonry veneer` type, and masonry veneer area Masonry veneer type has 24 NAs. Masonry veneer area has 23 NAs. If a house has a veneer area, it should also have a masonry veneer type. Let’s fix this one first

```{r}
#check if the 23 houses with veneer area NA are also NA in the veneer type
length(which(is.na(all$MasVnrType) & is.na(all$MasVnrArea)))
```
```{r}
#find the one that should have a MasVnrType
all[is.na(all$MasVnrType) & !is.na(all$MasVnrArea), c('MasVnrType', 'MasVnrArea')]
```
```{r}
#fix this veneer type by imputing the mode
all$MasVnrType[2611] <- names(sort(-table(all$MasVnrType)))[2] #taking the 2nd value as the 1st is 'none'
all[2611, c('MasVnrType', 'MasVnrArea')]
```

```{r}
Masonry <- c('None'=0, 'BrkCmn'=0, 'BrkFace'=1, 'Stone'=2)
all$MasVnrType<-as.integer(revalue(all$MasVnrType, Masonry))
table(all$MasVnrType)
```
MasVnrArea: Masonry veneer area in square feet.An integer variable.

```{r}
all$MasVnrArea[is.na(all$MasVnrArea)] <-0
```
  

`MSZoning`: Identifies the general zoning classification of the sale 4 NAs. Values are categorical.

   A    Agriculture
   C    Commercial
   FV   Floating Village Residential
   I    Industrial
   RH   Residential High Density
   RL   Residential Low Density
   RP   Residential Low Density Park 
   RM   Residential Medium Density
   
```{r}
#imputing the mode
all$MSZoning[is.na(all$MSZoning)] <- names(sort(-table(all$MSZoning)))[1]
all$MSZoning <- as.factor(all$MSZoning)
table(all$MSZoning)
```
Kitchen quality and numer of Kitchens above grade
`Kitchen quality` has 1 NA. Number of Kitchens is complete.
1NA. Can be made ordinal with the qualities vector.

   Ex   Excellent
   Gd   Good
   TA   Typical/Average
   Fa   Fair
   Po   Poor
   
```{r}
all$KitchenQual[is.na(all$KitchenQual)] <- 'TA' #replace with most common value
all$KitchenQual<-as.integer(revalue(all$KitchenQual, Qualities))
table(all$KitchenQual)
```
Number of Kitchens above grade. An integer variable with no NAs.
```{r}
table(all$KitchenAbvGr)
```
`Utilities`: Type of utilities available. 2 NAs. Ordinal as additional utilities is better.

   AllPub   All public Utilities (E,G,W,& S)    
   NoSewr   Electricity, Gas, and Water (Septic Tank)
   NoSeWa   Electricity and Gas Only
   ELO  Electricity only
   
However, the table below shows that only one house does not have all public utilities. This house is in the train set. Therefore, imputing ‘AllPub’ for the NAs means that all houses in the test set will have ‘AllPub’. This makes the variable useless for prediction. Consequently, We have get rid of it.
```{r}
table(all$Utilities)
```
```{r}

kable(all[is.na(all$Utilities) | all$Utilities=='NoSeWa', 1:9])
```

```{r}
all$Utilities <- NULL
```

`Functional: Home functionality`.1NA. Can be made ordinal (salvage only is worst, typical is best).
```{r}
#impute mode for the 1 NA
all$Functional[is.na(all$Functional)] <- names(sort(-table(all$Functional)))[1]

all$Functional <- as.integer(revalue(all$Functional, c('Sal'=0, 'Sev'=1, 'Maj2'=2, 'Maj1'=3, 'Mod'=4, 'Min2'=5, 'Min1'=6, 'Typ'=7)))
table(all$Functional)
```
`Exterior variables`
There are 4 exterior variables.2 variables have 1 NA, 2 variables have no NAs.
`Exterior1st`: Exterior covering on house. 1 NA. Values are categorical.

```{r}
#imputing mode
all$Exterior1st[is.na(all$Exterior1st)] <- names(sort(-table(all$Exterior1st)))[1]

all$Exterior1st <- as.factor(all$Exterior1st)
table(all$Exterior1st)
```
`Electrical: Electrical system`1 NA. Values are categorical.

   SBrkr    Standard Circuit Breakers & Romex
   FuseA    Fuse Box over 60 AMP and all Romex wiring (Average) 
   FuseF    60 AMP Fuse Box and mostly Romex wiring (Fair)
   FuseP    60 AMP Fuse Box and mostly knob & tube wiring (poor)
   Mix  Mixed
   
```{r}
#imputing mode
all$Electrical[is.na(all$Electrical)] <- names(sort(-table(all$Electrical)))[1]

all$Electrical <- as.factor(all$Electrical)
table(all$Electrical)
```
`SaleType`: Type of sale 1 NA. Values are categorical.

   WD   Warranty Deed - Conventional
   CWD  Warranty Deed - Cash
   VWD  Warranty Deed - VA Loan
   New  Home just constructed and sold
   COD  Court Officer Deed/Estate
   Con  Contract 15% Down payment regular terms
   ConLw    Contract Low Down payment and low interest
   ConLI    Contract Low Interest
   ConLD    Contract Low Down
   Oth  Other

```{r}
#imputing mode
all$SaleType[is.na(all$SaleType)] <- names(sort(-table(all$SaleType)))[1]

all$SaleType <- as.factor(all$SaleType)
table(all$SaleType)
```
`SaleCondition`: Condition of sale. No NAs. Values are categorical.

   Normal   Normal Sale
   Abnorml  Abnormal Sale -  trade, foreclosure, short sale
   AdjLand  Adjoining Land Purchase
   Alloca   Allocation - two linked properties with separate deeds, typically condo with a garage unit  
   Family   Sale between family members
   Partial  Home was not completed when last assessed (associated with New Homes)

```{r}
all$SaleCondition <- as.factor(all$SaleCondition)
table(all$SaleCondition)
```
## 4.3 Label encoding/factorizing the remaining character variables.
At this point, We have made sure that all variables with NAs are taken care of. However, We still need to also take care of the remaining character variables that without missing values. Similar to the previous section, We have created Tabs for groups of variables
```{r}
Charcol <- names(all[,sapply(all, is.character)])
Charcol
```

```{r}
#No ordinality, so converting into factors
all[c("Foundation", "Heating", "RoofStyle","RoofMatl", "BldgType","HouseStyle", "LandContour", "Neighborhood", "Condition1", "Condition2", "OverallQual")] <- 
  lapply(all[c("Foundation", "Heating", "RoofStyle", "RoofMatl", "BldgType","HouseStyle", "LandContour", "Neighborhood", "Condition1", "Condition2","OverallQual")], as.factor)
```


```{r}
#Ordinal, so label encoding
all$LandSlope<-as.integer(revalue(all$LandSlope, c('Sev'=0, 'Mod'=1, 'Gtl'=2)))
all$CentralAir<-as.integer(revalue(all$CentralAir, c('N'=0, 'Y'=1)))
all$HeatingQC<-as.integer(revalue(all$HeatingQC, Qualities))
all$Street<-as.integer(revalue(all$Street, c('Grvl'=0, 'Pave'=1)))
all$PavedDrive<-as.integer(revalue(all$PavedDrive, c('N'=0, 'P'=1, 'Y'=2)))
```

## 4.4 Changing some numeric variables into factors.
At this point, most of the variables are complete (No NAs), and all character variables are converted into either numeric labels or into factors. However, there are 3 variables that are recorded numeric but should actually be categorical.
### 4.4.1 Year and Month Sold
While oridinality within YearBuilt (or remodeled) makes sense (old houses are worth less), we are talking about only 5 years of sales. Month Sold is also an Integer variable. However, December is not “better” than January. Therefore, I will convert MoSold values back into factors.
```{r}
all$MoSold <- as.factor(all$MoSold)
```

### 4.4.2 MSSubClass
`MSSubClass`: Identifies the type of dwelling involved in the sale.
    20  1-STORY 1946 & NEWER ALL STYLES
    30  1-STORY 1945 & OLDER
    40  1-STORY W/FINISHED ATTIC ALL AGES
    45  1-1/2 STORY - UNFINISHED ALL AGES
    50  1-1/2 STORY FINISHED ALL AGES
    60  2-STORY 1946 & NEWER
    70  2-STORY 1945 & OLDER
    75  2-1/2 STORY ALL AGES
    80  SPLIT OR MULTI-LEVEL
    85  SPLIT FOYER
    90  DUPLEX - ALL STYLES AND AGES
   120  1-STORY PUD (Planned Unit Development) - 1946 & NEWER
   150  1-1/2 STORY PUD - ALL AGES
   160  2-STORY PUD - 1946 & NEWER
   180  PUD - MULTILEVEL - INCL SPLIT LEV/FOYER
   190  2 FAMILY CONVERSION - ALL STYLES AND AGES
These classes are coded as numbers, but really are categories.
```{r}
str(all$MSSubClass)
```
```{r}
all$MSSubClass <- as.factor(all$MSSubClass)
```

## 5. Visualization of important variables
```{r}
numericVars <- which(sapply(all, is.numeric)) #index vector numeric variables
factorVars <- which(sapply(all, is.factor)) #index vector factor variables
cat('There are', length(numericVars), 'numeric variables, and', length(factorVars), 'categoric variables')
```

## Total number of Bathrooms
There are 4 bathroom variables. Individually, these variables are not very important. However, I assume that I if I add them up into one predictor, this predictor is likely to become a strong one.
“A half-bath, also known as a powder room or guest bath, has only two of the four main bathroom components-typically a toilet and sink.” Consequently, I will also count the half bathrooms as half.
```{r}
all$TotBathrooms <- all$FullBath + (all$HalfBath*0.5) + all$BsmtFullBath + (all$BsmtHalfBath*0.5)
```

```{r}
# graphical justification
library(gridExtra)
tb1 <- ggplot(data=all[!is.na(all$SalePrice),], aes(x=as.factor(TotBathrooms), y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
tb2 <- ggplot(data=all, aes(x=as.factor(TotBathrooms))) +
        geom_histogram(stat='count')
grid.arrange(tb1, tb2)
```
### Total Square Feet
As the total living space generally is very important when people buy houses, I am adding a predictors that adds up the living space above and below ground.
```{r}
all$TotalSqFeet <- all$GrLivArea + all$TotalBsmtSF
```

```{r}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=TotalSqFeet, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_text_repel(aes(label = ifelse(all$GrLivArea[!is.na(all$SalePrice)]>4500, rownames(all), '')))
```
```{r}
cor(all$SalePrice, all$TotalSqFeet, use= "pairwise.complete.obs")
```
As expected, the correlation with SalePrice is very strong indeed (0.78)
The two potential outliers seem to ‘outlie’ even more than before. By taking out these two outliers, the correlation increases by 5%.

```{r}
cor(all$SalePrice[-c(524, 1299)], all$TotalSqFeet[-c(524, 1299)], use= "pairwise.complete.obs")
```

## 7.1 Dropping highly correlated variables

```{r}
#dropVars <- c('YearRemodAdd', 'GarageYrBlt', 'GarageArea', 'GarageCond', 'TotalBsmtSF', 'TotalRmsAbvGrd', 'BsmtFinSF1')

#all <- all[,!(names(all) %in% dropVars)]
```

## 7.2 Removing outliers

```{r}
all <- all[-c(524, 1299),]
```



```{r}
train <- all[!is.na(all$SalePrice),]
test <- all[is.na(all$SalePrice),]
train
train %>% filter(GrLivArea<4500)
```



## 6. Cross validation and Modeling

Cross-validation (CV) a method to avoid overfitting by estimating the model's performance with new data.  The simplest CV technique is the validation set method, which is baked into Kaggle's structure:  the validation set is held back for model evaluation.  To estimate how our model will perform at Kaggle, however, we need to create a train and validation set *using* the train set.

Steps.

1. Split train set into two "folds" (train fold and validation fold).  A 70/30 split is common.
2. Fit the model to the train fold and evaluate it on the validation fold.  RMSE and R-squared on the validation fold will be the estimate of the model's out of sample performance.

```{r}
?sample
# Randomly sample 70% of the rows
set.seed(115)
index <- sample(x = 1:nrow(train), size = nrow(train)*.7, replace = F)

head(index) # These are row numbers
```

```{r}
# Subset train using the index to create train_fold
train_fold <- train[index, ]

# Subset the remaining row to create validation fold.
validation_fold <- train[-index, ]
```

Now fit model with the `train_fold` and test it on the `validation_fold`:

```{r}

# Fit example model
#model <- lm(SalePrice ~ GarageCars + OverallQual + MSSubClass + GrLivArea + Neighborhood + MSZoning + PoolQC +LotFrontage + BsmtQual  + BsmtExposure + BsmtFinType1 +BsmtUnfSF + KitchenQual + Functional + Foundation  + HeatingQC  + Condition1 + TotBathrooms + TotalSqFeet + CentralAir + YearBuilt + ExterQual +GarageQual + LotShape+ LotArea+ LotConfig + Alley + Fireplaces + FireplaceQu , data = train_fold)

model <- lm(SalePrice ~ OverallQual*GrLivArea + GrLivArea * YearBuilt + GrLivArea * YearRemodAdd + GarageCars + GarageArea + TotalBsmtSF*`1stFlrSF` + GrLivArea*`2ndFlrSF` + GrLivArea*FullBath + GrLivArea*TotalBsmtSF + GrLivArea * TotRmsAbvGrd + GrLivArea*BsmtFinSF1 + OverallQual*BsmtFinSF2 + TotalBsmtSF*Neighborhood + GrLivArea*MasVnrArea + Fireplaces + GarageYrBlt + TotalBsmtSF*LotArea + YearBuilt*Condition1 + OverallQual*HouseStyle + GrLivArea * BldgType + OverallCond * HalfBath + TotalBsmtSF*SaleCondition - KitchenAbvGr, data = train_fold) 


# Get predictions for the validation fold
predictions <- predict(model, newdata = validation_fold)

# Create functions for calculating RMSE and R-squared (necessary for estimating 
# out of sample performance)

rmse <- function(observed, predicted) sqrt(mean((observed - predicted)^2))

R2 <- function(observed, predicted){
  TSS <- sum((observed - mean(observed))^2)
  RSS <- sum((observed - predicted)^2)
  1- RSS/TSS
}

rmse(validation_fold$SalePrice, predictions)
R2(validation_fold$SalePrice, predictions)

```
## 7.Test set modeling.

1. Fit the model using the entire train set.
2. Make exactly the same changes to the test set that you made to the train set.
3. Check there are no missing observations for your selected predictors in the test set.
4. Make predictions for the test set.
5. Format your submission file.
6. Submit to Kaggle.

```{r}
# 1. Fitting model to the entire train set.
submission_model <- lm(SalePrice ~ OverallQual*GrLivArea + GrLivArea * YearBuilt + GrLivArea * YearRemodAdd + GarageCars + GarageArea + TotalBsmtSF*`1stFlrSF` + GrLivArea*`2ndFlrSF` + GrLivArea*FullBath + GrLivArea*TotalBsmtSF + GrLivArea * TotRmsAbvGrd + GrLivArea*BsmtFinSF1 + OverallQual*BsmtFinSF2 + TotalBsmtSF*Neighborhood + GrLivArea*MasVnrArea + Fireplaces + GarageYrBlt + TotalBsmtSF*LotArea + YearBuilt*Condition1 + OverallQual*HouseStyle + GrLivArea * BldgType + OverallCond * HalfBath + TotalBsmtSF*SaleCondition - KitchenAbvGr, data = train)
summary(submission_model)
```

## imputing different level in test data for MSSubClass
For MSSubclass there seems to be different levels in the test set which are not present in the train set
```{r}
#levels on train data
levels(train$MSSubClass)
```
```{r}
#levels on test data
levels(test$MSSubClass)

```
We can observe from the above display of levels there is an additional level of 150 added to the test set.
from the  90	DUPLEX - ALL STYLES AND AGES
          120	1-STORY PUD (Planned Unit Development) - 1946 & NEWER
          150	1-1/2 STORY PUD - ALL AGES 
I am replacing 150 with 90 rather than zero to match the nearest possible level details.(i.e ALL AGES)
```{r}
if ("150" %in% levels(test$MSSubClass)) {
  test$MSSubClass <- as.character(test$MSSubClass)  # Convert to character for replacement
  test$MSSubClass[test$MSSubClass == "150"] <- "90"
  test$MSSubClass <- as.factor(test$MSSubClass)  # Convert back to factor
}
levels(test$MSSubClass)
```


```{r}
# 4. Make predictions for the test set.
submission_predictions <- predict(submission_model, newdata = test) # Use the newdata argument!

head(submission_predictions)
```
# 5. Format your submission file.

reformatting test set with its id's
```{r}
test$Id <- test_labels
```


```{r}
submission <- test %>% 
  select(Id) %>% 
  mutate(SalePrice = submission_predictions)

# Check
head(submission)

# write to csv
write.csv(submission, "kaggle_submission2.csv", row.names = F)

```


We have the predicted values of SalePrice for houses in Ames, IA using multi linear regression modeling.

## 9.Score from Kaggle!


